---
title: "R Notebook"
output: html_notebook
---
Encoding categorical variables

What is a categorical variable?
In layman's terms, a quantity that is not a number but rather is a string or something of that sort is a categorical variable.


Why do we need to encode the categorical variables?

Usually in machine learning, the data fed to the model or the results calculated are numeric.


For example, say we want to predict a crop for a farmer to grow. Let the inputs be ambient temperature, average rainfall and soil type (red, black, loamy, sandy) in his farm land. The output would be the name of a crop.

A problem arises.

Not all inputs are in numeric format. The soil types are of string type. Neither is the output numeric.

Thus we have to encode this data to convert it into numerical values.



```{r}
# Importing dataset

data = read.csv('Data.csv')

print(data)
```

There are categorical variables, and also missing data.

We will first handle the missing data.

```{r}
# We will fill these NA values by the mean of the columns

data$Age = ifelse(is.na(data$Age),
                ave(data$Age, FUN =  function(x) mean(x, na.rm = TRUE)),
                data$Age)

data$Salary = ifelse(is.na(data$Salary),
                  ave(data$Salary, FUN =  function(x) mean(x, na.rm = TRUE)),
                  data$Salary)

print(data)
```

Now we encode the categorical variables, using the factor function.

```{r}
# factor(x = character(), levels, labels = levels, exclude = NA, ordered = is.ordered(x), nmax = NA)
# params:
# 1. x: a vector of data
# 2. levels: optional vector of unique values that x might have taken
# 3. labels: optional character vector of labels for levels (in same order as levels after removing those in exclude), or character string of length 1
# 4. exclude: vector of values to be excluded when forming set of levels
# 5. ordered: logical flag to determine if the levels should be regarded as ordered (in order given)
# 6. nmax: an upper bound on number of levels

data$Country = factor(data$Country,
                      levels = c('France', 'Spain', 'Germany'),
                      labels = c(0, 1, 2))

print(data)
```

This method is known as label encoding. Thus we have implemented label encoding of categorical variables in R
