---
title: "R Notebook"
output: html_notebook
---

Install the required library if not installed before

```{r}
# install.packages("mlbench")
```

This library is used to load inbuilt datasets, much like the sklearn.datasets library in Python

Load the library

```{r}
library(mlbench)

# list contents of the library

library(help = 'mlbench')
```

Load the boston housing dataset

```{r}
data("BostonHousing")
dim("BostonHousing")
head("BostonHousing")

```

We can see tha the library has 14 cols, 13 independent variables, 1 dependent variable.

To split the dataset into training and test set we are going to use the library "caTools".
```{r}
library(caTools)
```

Split data from vector Y into two sets in predefined ratio while preserving relative ratios of different labels in Y. 
Used to split the data used during classification into train and test subsets.
```{r}
split = sample.split(BostonHousing$medv, SplitRatio = 0.8)

training_set = subset(BostonHousing, split == TRUE)
test_set = subset(BostonHousing, split == FALSE)

```
Fitting Multiple Linear Regression to the Training set

```{r}
regressor = lm(formula = medv ~ .,
               data = training_set)
```
Predicting the Test Set results

```{r}
y_pred = predict(regressor, newdata = test_set)
```
Create the RMSE function, and print the RMSE value
```{r}

RMSE = function(m, o){
  sqrt(mean((m - o)^2))
}

print(RMSE(test_set$medv, y_pred))


```

